Corporate Security Solutions for BYOD:\\ A Novel User-Centric and Self-Adaptive System

### 2015-07-09 Response to reviewers - r2.tex

% Es quotation, no verbatim, øno?
% Antonio - con quotation no se distinguÌa bien el texto del revisor y el de la respuesta. SÛlo hacÌa un sangrado. Creo que asÌ se ve mejor.

\begin{verbatim}
After reviewing the new version of the document, I admit the authors
have  substantially changed section 3 (state of the art) and they have
improved and extended the introduction and sections 4 (description of
the system) and section 5 (comparison with other systems) I
acknowledge the ideas are really interesting and exciting, but I'm
still not sure if the authors are providing new results not published
in their previous works or just rephrasing their text.

Since there is an obvious over-the-top quality in their work and some
good ideas, I'll let the editors decide if the contents are novel
enough (compared to other previous works by the same authors) to be
published in the journal.
\end{verbatim}

We would like to thank the reviewer for his/her opinion and would lead him/her to have another look at the new version, since it has been highly improved.  
As a summary:
%
\begin{itemize}
\item We have described in a more detailed way our previously published papers in this scope (in Section 1), clarifying the differences and contributions of this work with respect to them.
\item Section 2 has been deeply improved adding a much more complete state of the art in the scope of this work. To this end, several research works have been read and commented in that section and along the text. This has helped us to better compare our system against the other products available in the market. Also, an analysis of the best practices for BYOD have been applied for each product and our system.
% Antonio - TODO: Fergu, completa esto t˙ que tienes la idea global de lo nuevo en el SotA, por favor. FERGU: Hecho. Metido tambiÈn lo de las best practices
New references added after the review can be found at the end of this letter.

\item Section 3 has been rewritten in a big part, including a new taxonomy for the analyzed systems at the beginning of the section. This taxonomy has been created taking into account the expertise extracted from the new references of the state of the art, such as the best practices by Romer \cite{Romer14BestPractices} or the hands-off/hands-on classification by Scarfo et al. \cite{Scarfo12survey}, among others. Furthermore, some new tools have been added, and all the products have been better classified considering the defined taxonomy.
% Antonio - TODO: extender esto de la secciÛn 3 y 5 un poco m·s. FERGU: Done

@@

% -----------------------------------------------------------------------------

\section{Comments by Reviewer \#3}

\begin{verbatim}
This paper describes MUSES, an adaptable framework for enforcing 
security policies on mobile devices. This tool relies on Machine 
Learning and CI techniques to predict policy violations and is 
self-adaptive on user behaviors.
The idea seems interesting although several technical concerns 
arise from the reading of this work:
1.      The tools aims at enforcing security policies, however 
the security response occurs only AFTER a particular pattern is 
recognized to be potentially malicious. From my understandings, 
there is no a-priori enforcement except for the prediction based 
on event logs. This could mean that a security violation could 
in principle happen and only later recognized. This is not so 
suitable for a corporate environment with strict security 
requirements. This statement should be discussed and preferably 
discredited with experimental results.
\end{verbatim}

MUSES starts from a defined set of security rules that the CSO must
define. Ideally, these rules should deal, in principle, with all the
potential security incidents that the company aims to detect. MUSES
would be able to define new rules, by means of refinement, which could
be labeled, i.e. assigned a decision, after the corresponding events
have happened. However, it can also infer or create new rules, which
the CSO must approve, using computational intelligence
techniques. These rules could deal with unexpected situations, which
have not happened previously .

@@

Anyway, the idea is to define a set of `restrictive' rules at the very beginning, able to ensure the system safety in a big amount of situations. Then, MUSES will refine or improve this set, making more specific and optimal rules.

Moreover, MUSES offers visual analysis tools to the CSOs, which can
aid or guide them to detect suspicious or anomalous patterns, and to
study how to avoid these patterns becoming malicious to the company.

In order to clarify this we have rewritten and completed the second
paragraph in section 4.1 (page XX). It is now:

\textit{The client is an important part of the a-priori risk treatment, because it both monitors the whole context, and finally acts or warns the user in event of a security violation. 
But security violations are mainly identified by security rules, which is why MUSES needs at least an initial set of rules, defined by the CSO. Then, it makes use of DM and CI techniques to improve or optimize them, enhancing the coverage. 
That is to say, the more sensors are implemented, the more information related to event context can be extracted. Implementing a sensor means to include a programmed method which triggers MUSES to start `listening' to some parts of a device. Then, the initial security rules establish which situations might cause a security violation, and the sensors provide with information which helps the MUSES server to identify events that are alike to the ones that are already known as dangerous.}

@@

We have also improved this clarification in section 5 (page XX), adding the text:

\textit{However, MUSES presents a limitation regarding the enhancement of rules, since, in principle, it cannot predict (generate rules for dealing with) unexpected or unknown events which could lead to a security incident. The philosophy is that the initial set of rules, defined by the CSO, should be very restrictive regarding possible unexpected users' behaviors or events, in order to avoid as much security incidents as possible. As the system works, MUSES would be able to define new rules (through refinement) which could get an associated decision after the corresponding events have happened. Thus, this will lead to get eventually an optimal set of security rules.}

@@

The abstract has been changed and that confusing sentence has been
eliminated. Now we think it is clear, as intended, the motivation,
context and focus of the paper.

\begin{verbatim}
The description of IBM products, in section 3.1, does not provide
clear distinction between various service types provided by IBM. It
furthermore lacks focused coverage of services relevant for this
work.

In analysis of IBM offer, authors refer to general web site for mobile
security products and consumer guide/whitepaper that explores the
general problem of BYOD. They consider services that are not purely
related to BYOD and MUSES. For instance, it is not clear how
"Enterprise Wireless Networks" (page 6, lines 21-26) fit into
comparison, and how establishment of secure connections can be related
to MUSES's adaptability. On the other hand, the products such as
MobileFirst and MaaS360 are omitted. The suggestion in this regard is
to consider rewriting this part, including updated references and
current tools in comparison. 
\end{verbatim}

Following reviewer's suggestions, this section has been rewritten, and
the references mentioned by the referee, including new products, have
been considered in the study. The different IBM tools and services have been described 
(end of Section 3.1, in page XX), and compared to MUSES in Section 5.
%FERGU: PONED ESTO M√ÅS BONICO!
% Antonio - ahora est· un poco mejor, pero no sÈ si m·s bonico. Fergu debes justificar el por quÈ se han puesto los tres servicios de IBM, ya que el revisor dice que algunos no est·n relacionados con el tema del trabajo... t˙ que sabes de esto de IBM. ;D
% Fergu, tambiÈn falta la comparativa con MUSES de esas herramientas (SecciÛn 5)...
% [Pedro] he cambiado un poco la redacciÛn   ;)
% FERGU: Bueno, al final estos productos forman parte del "Entorno IBM" en general, que sÌ se ha comparado con MUSES. Paloma ha reescrito tambiÈn la secciÛn y quitado lo que sobraba.

@@

1) Explain why particular solutions are chosen for inclusion in review 
and what is their relevance? Did authors consider other (possible) 
solutions relying on DM and CI?
\end{verbatim}

\textcolor{blue}{TODO}

\begin{verbatim}
 2) Considering variety of BYOD tools and their different features, 
authors might consider to introduce or reuse a taxonomy or 
classification of those approaches, and state the place of MUSES 
there (or define a new subcategory, if deemed necessary)
\end{verbatim}

Following the reviewer's suggestion we have introduced a taxonomy for BYOD, according to the works that we have revised in the state of the art (Section 2). Different concepts mentioned in this improved state of the art, such as the hands-off/hand-off differentiation or BYOD best practices, have been taken into account to justify the proposed taxonomy. This taxonomy is described at the beginning of Section 3 (second paragraph, page XX). Then, all the tools, including MUSES, have been classified in the types defined by the taxonomy, and summarized in Table 1.
% Antonio - completar con el n˙mero de p·gina final
% Antonio - TODO: decir si MUSES y el resto de sistemas se han encuadrado dentro de esa taxonomÌa
% FERGU: completado lo de la taxonomia (falta lo de los n˙meros de p·gina a˙n!) TODO

@@

\begin{verbatim}
2) Authors overuse parentheses to explain details. This hinders 
readability and text flow. The level of details provided and their 
relevance in the context are only partially consistent. On p.14 
users introduce footnotes for additional details. Maybe they 
should consider applying such approach in other parts of the 
document as well.
\end{verbatim}

Following reviewer's suggestion we have rewritten the majority 
of the phrases where parentheses were used, avoiding them for 
the sake of a better text flow and reading.

\begin{verbatim}
3) The structure and flow of the work can be improved.
\end{verbatim}

We have rewritten several parts of the paper and included new sections, subsections, tables and lists in order to improve the structure, contents and fluency of the text. Moreover, some paragraphs have been also rewritten to avoid large phrases, the inadequate use of parentheses, or the misunderstanding of some concepts.
% Antonio - Por favor, completad esta justificaciÛn mejor.

### 2015-07-09 MUSES_vs_BYOD_solutions.tex
Lista como descripciones

% Antonio - pongo esto como lista de items, ya que era la única lista diferente
\begin{itemize}
  \item \textit{Connectivity sensor:} It gathers all information related to network communications, in addition to security properties like network encryption, or number of neighbours. This allows MUSES actuator to forbid the user for sending important emails, or opening key assets if, for instance, the network encryption is not secure enough. 
  \item \textit{Device protection sensor:} Which scans device security settings, such as the time until the device screen turns off, if the device has the screen protected with a password, or if the device is rooted \cite{hildenbrand2014}. This security mechanism adds the possibility of detecting security violations even if the user is not doing a specific action.
  \item \textit{Location sensor:} User location can be obtained using the device GPS. Although, this might be taken against the preservation of user privacy. This is why MUSES allows the definition of a number of `zones', and therefore it only checks if the user is inside them or not, acting in consequence.
  \item \textit{Mail sensor:} If the user is writing an email, MUSES client sensor can also check the email attributes. Thus, it can detect if the event context is dangerous, according to security policies and the RT2AE, before the mail is sent.
  % (Paloma) Pedro, aquí he cambiado el "is going to send" a "is writing" en vez de "is sending" porque en realidad lo que queremos es que se le avise antes de mandarlo, no cuando lo está mandando o lo ha mandado. Se explica después. ¿Cómo lo veis?
  \item \textit{File sensor:} The type of file, also called \textit{asset}, is also very important for the application of security rules. Thus, this sensor aims to detect if some files might need more strict security policies, also depending on the rest of sensors.
  \item \textit{App sensor:} It is in charge of monitoring the information about the application from which the user is performing the action, but also looks for installed trusted antivirus, and encryption applications. Additionally, this sensor gathers information about the applications which the user tries to install or uninstall.
\end{itemize}

>>>>

\begin{description}
  \item[Connectivity sensor:] It gathers all information related to network communications, in addition to security properties like network encryption, or number of neighbours. This allows MUSES actuator to forbid the user for sending important emails, or opening key assets if, for instance, the network encryption is not secure enough. 
  \item[Device protection sensor:] Which scans device security settings, such as the time until the device screen turns off, if the device has the screen protected with a password, or if the device is rooted \cite{hildenbrand2014}. This security mechanism adds the possibility of detecting security violations even if the user is not doing a specific action.
  \item[Location sensor:] User location can be obtained using the device GPS. Although, this might be taken against the preservation of user privacy. This is why MUSES allows the definition of a number of `zones', and therefore it only checks if the user is inside them or not, acting in consequence.
  \item[Mail sensor:] If the user is writing an email, MUSES client sensor can also check the email attributes. Thus, it can detect if the event context is dangerous, according to security policies and the RT2AE, before the mail is sent.
  % (Paloma) Pedro, aquí he cambiado el "is going to send" a "is writing" en vez de "is sending" porque en realidad lo que queremos es que se le avise antes de mandarlo, no cuando lo está mandando o lo ha mandado. Se explica después. ¿Cómo lo veis?
  \item[File sensor:] The type of file, also called \textit{asset}, is also very important for the application of security rules. Thus, this sensor aims to detect if some files might need more strict security policies, also depending on the rest of sensors.
  \item[App sensor:] It is in charge of monitoring the information about the application from which the user is performing the action, but also looks for installed trusted antivirus, and encryption applications. Additionally, this sensor gathers information about the applications which the user tries to install or uninstall.
\end{description}
